---
title: "Hands-on Exercise 1: Geospatial Data Wrangling with R"
execute:
  warning: false
date: "`r Sys.Date()`"
---

# 1 Getting Started

## 1.1 Installing and loading R packages

In this section, I will install and load tidyverse and sf packages.

```{r}
pacman::p_load(tidyverse, sf)
```

## 1.2 Importing a geospatial data

### 1.2.1 Importing polygon feature data in shapefile format

```{r}
mpsz <- st_read(dsn = "data/geospatial", layer = "MP14_SUBZONE_WEB_PL" )

```

### 1.2.2 Importing polyine feature data in shapefile form
```{r}
cyclingpath = st_read(dsn = "data/geospatial", layer = "CyclingPathGazette")
```

### 1.2.3 Importing GIS data in kml format
```{r}
preschool = st_read("data/geospatial/preschools-location.kml")
```

## 1.3 Checking the content of a simple feature data frame
### 1.3.1 Working with st_geometry()
```{r}
st_geometry(mpsz)
```

### 1.3.2 Working with glimpse()
```{r}
glimpse(mpsz)
```

### 1.3.3 Working with head()
```{r}
head(mpsz, n=5)
```

## 1.4 Plotting the Geospatial Data

```{r}
plot(mpsz)

```
