---
title: "In-class Exercise 4 - Hands on Exercise 5"
author: "Kwek Ming Rong"
date: "30 January 2023"
date-modified: "`r Sys.Date()`"
format: html
execute:
  eval: true
  echo: true
  warning: false
editor: visual
---

# 5. Getting Started - 2nd Order Spatial Point Patterns Analysis

## 5.1 Overview

Spatial Point Pattern Analysis is the evaluation of the pattern or distribution, of a set of points on a surface. The point can be location of:

-   events such as crime, traffic accident and disease onset, or

-   business services (coffee and fastfood outlets) or facilities such as childcare and eldercare.

Using appropriate functions of [spatstat](https://cran.r-project.org/web/packages/spatstat/), this hands-on exercise aims to discover the spatial point processes of childecare centres in Singapore.

The specific questions we would like to answer are as follows:

-   are the childcare centres in Singapore randomly distributed throughout the country?\

-   if the answer is not, then the next logical question is where are the locations with higher concentration of childcare centres?

## 5.2 The data

To provide answers to the questions above, three data sets will be used. They are:

-   `CHILDCARE`, a point feature data providing both location and attribute information of childcare centres. It was downloaded from Data.gov.sg and is in geojson format.

-   `MP14_SUBZONE_WEB_PL`, a polygon feature data providing information of URA 2014 Master Plan Planning Subzone boundary data. It is in ESRI shapefile format. This data set was also downloaded from Data.gov.sg.

-   `CostalOutline`, a polygon feature data showing the national boundary of Singapore. It is provided by SLA and is in ESRI shapefile format.

## 5.3 Installing and Loading the R packages

In this hands-on exercise, five R packages will be used, they are:

-   [**sf**](https://r-spatial.github.io/sf/), a relatively new R package specially designed to import, manage and process vector-based geospatial data in R.

-   [**spatstat**](https://spatstat.org/), which has a wide range of useful functions for point pattern analysis. In this hands-on exercise, it will be used to perform 1st- and 2nd-order spatial point patterns analysis and derive kernel density estimation (KDE) layer.

-   [**raster**](https://cran.r-project.org/web/packages/raster/) which reads, writes, manipulates, analyses and model of gridded spatial data (i.e. raster). In this hands-on exercise, it will be used to convert image output generate by spatstat into raster format.

-   [**maptools**](https://cran.r-project.org/web/packages/maptools/) which provides a set of tools for manipulating geographic data. In this hands-on exercise, we mainly use it to convert *Spatial* objects into *ppp* format of **spatstat**.

-   [**tmap**](https://cran.r-project.org/web/packages/tmap/) which provides functions for plotting cartographic quality static point patterns maps or interactive maps by using [leaflet](https://leafletjs.com/) API.

Use the code chunk below to install and launch the five R packages.

```{r}
pacman::p_load(maptools, sf, raster, spatstat, tmap)
```

## 5.4 Spatial Data Wrangling - Importing Data

Things to learn from this code chunk.

```{r}
childcare_sf <- st_read("data/Geospatial/childcare.geojson") %>%
  st_transform(crs = 3414)
```

```{r}
sg_sf <- st_read(dsn = "data/Geospatial", layer="CostalOutline")
```

```{r}
mpsz_sf <- st_read(dsn = "data/geospatial", 
                layer = "MP14_SUBZONE_WEB_PL")
```

## 5.5 Converting sf data frames to sp's Spatial\* class

The code chunk below uses as_Spatial() of sf package to convert the three geospatial data from simple feature data frame to sp's Spatial\* class.

```{r}
childcare <- as_Spatial(childcare_sf)
mpsz <- as_Spatial(mpsz_sf)
sg <- as_Spatial(sg_sf)
```

Plot the childcare centre in Singapore (Visualising the sf layers)

```{r}
sf_use_s2(FALSE)
tmap_mode('view')
tm_shape(childcare_sf) +
  tm_dots(alph = 0.5, 
          size = 0.01) +
  tm_view(set.zoom.limits = c(11,14)) ## 11 is zoom out value, 14 zoom in value
```
## 5.6 Converting the Spatial\* class into generic sp format

spatstat requires the analytical data in ppp object form. There is no direct way to convert a Spatial\* classes into ppp object. We need to convert the Spatial classes\* into Spatial object first.

The codes chunk below converts the Spatial\* classes into generic sp objects dropping the table from the dataframe.

```{r}
childcare_sp <- as(childcare, "SpatialPoints")
sg_sp <- as(sg, "SpatialPolygons")
```

## 5.7 Converting the generic sp format into spatstat's ppp format

Now, we will use as.ppp() function of spatstat to convert the spatial data into spatstat's ppp object format.

```{r}
childcare_ppp <- as(childcare_sp, "ppp")
childcare_ppp
```

## 5.8 Handling duplicated points

```{r}
any(duplicated(childcare_ppp))
multiplicity(childcare_ppp)
sum(multiplicity(childcare_ppp) > 1)
```

Jittering approach 

```{r}
childcare_ppp_jit <- rjitter(childcare_ppp, 
                             retry=TRUE, 
                             nsim=1, 
                             drop=TRUE)

any(duplicated(childcare_ppp_jit))
```

## 5.9 Creating owin object

```{r}
sg_owin <- as(sg_sp, "owin")
```

## 5.10 Combining points events object and owin object

```{r}
childcareSG_ppp = childcare_ppp[sg_owin]
summary(childcareSG_ppp)
```

Extracting the study area

```{r}
pg = mpsz[mpsz@data$PLN_AREA_N == "PUNGGOL",]
tm = mpsz[mpsz@data$PLN_AREA_N == "TAMPINES",]
ck = mpsz[mpsz@data$PLN_AREA_N == "CHOA CHU KANG",]
jw = mpsz[mpsz@data$PLN_AREA_N == "JURONG WEST",]
```

```{r}
par(mfrow=c(2,2))
plot(pg, main = "Punggol")
plot(tm, main = "Tampines")
plot(ck, main = "Choa Chu Kang")
plot(jw, main = "Jurong West")
```

Converting the spatial point data frame into generic sp format

```{r}
pg_sp = as(pg, "SpatialPolygons")
tm_sp = as(tm, "SpatialPolygons")
ck_sp = as(ck, "SpatialPolygons")
jw_sp = as(jw, "SpatialPolygons")
```

Creating owin object

```{r}
pg_owin = as(pg_sp, "owin")
tm_owin = as(tm_sp, "owin")
ck_owin = as(ck_sp, "owin")
jw_owin = as(jw_sp, "owin")
```

Combining childcare points and the study area

```{r}
childcare_pg_ppp = childcare_ppp_jit[pg_owin]
childcare_tm_ppp = childcare_ppp_jit[tm_owin]
childcare_ck_ppp = childcare_ppp_jit[ck_owin]
childcare_jw_ppp = childcare_ppp_jit[jw_owin]
```


Converting from metre to kilometre
```{r}
childcare_pg_ppp.km = rescale(childcare_pg_ppp, 1000, "km")
childcare_tm_ppp.km = rescale(childcare_tm_ppp, 1000, "km")
childcare_ck_ppp.km = rescale(childcare_ck_ppp, 1000, "km")
childcare_jw_ppp.km = rescale(childcare_jw_ppp, 1000, "km")
```

```{r}
par(mfrow=c(2,2))
plot(childcare_pg_ppp.km, main="Punggol")
plot(childcare_tm_ppp.km, main="Tampines")
plot(childcare_ck_ppp.km, main="Choa Chu Kang")
plot(childcare_jw_ppp.km, main="Jurong West")
```

## 5.11 Second order spatial point pattern analysis

### 5.11.1 Analysing Spatial Point Process using G Function

The G function measures the distribution of the distances from an arbitrary event to its nearest event. In this section, you will learn how to compute G-function estimation by using Gest() of spatstat package. You will also learn how to perform monta carlo simulation test using envelope() of spatstat package.

Choa Chu Kang planning area
```{r}
G_CK = Gest(childcare_ck_ppp, correction = "border")
plot(G_CK, xlim=c(0,500))
```

Performing Complete Spatial Randomness Test

```{r}
G_CK.csr <- envelope(childcare_ck_ppp, Gest, nsim = 999)
```

```{r}
plot(G_CK.csr)
```

Tampines planning area

```{r}
G_tm = Gest(childcare_tm_ppp, correction = "best")
plot(G_tm)
```

Performing Complete Spatial Randomness Test

```{r}
G_tm.csr <- envelope(childcare_tm_ppp, Gest, correction = "all", nsim = 999)
```

```{r}
plot(G_tm.csr)
```

### 5.11.2 Analysing Spatial Point Process using F Function

Choa Chu Kang planning area
```{r}
F_CK = Fest(childcare_ck_ppp)
plot(F_CK)
```

Performing Complete Spatial Randomness Test

```{r}
F_CK.csr <- envelope(childcare_ck_ppp, Fest, nsim = 999)
```

```{r}
plot(F_CK.csr)
```

Tampines planning area

```{r}
F_tm = Fest(childcare_tm_ppp, correction = "best")
plot(F_tm)
```


```{r}
F_tm.csr <- envelope(childcare_tm_ppp, Fest, correction = "all", nsim = 999)
```


```{r}
plot(F_tm.csr)
```

### 5.11.3 Analysing Spatial Point Process using K Function

Choa Chu Kang planning area

```{r}
K_ck = Kest(childcare_ck_ppp, correction = "Ripley")
plot(K_ck, . -r ~ r, ylab= "K(d)-r", xlab = "d(m)")
```

Performing Complete Spatial Randomness Test

```{r}
K_ck.csr <- envelope(childcare_ck_ppp, Kest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(K_ck.csr, . - r ~ r, xlab="d", ylab="K(d)-r")
```

Tampines Planning Area

```{r}
K_tm = Kest(childcare_tm_ppp, correction = "Ripley")
plot(K_tm, . -r ~ r, 
     ylab= "K(d)-r", xlab = "d(m)", 
     xlim=c(0,1000))
```

```{r}
K_tm.csr <- envelope(childcare_tm_ppp, Kest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r}
plot(K_tm.csr, . - r ~ r, 
     xlab="d", ylab="K(d)-r", xlim=c(0,500))
```

### 5.11.4 Analysing Spaitial Point Process using L Function

Choa Chu Kang Planning Area

```{r}
L_ck = Lest(childcare_ck_ppp, correction = "Ripley")
plot(L_ck, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)")
```

Performing Complete Spatial Randomness Test

```{r}
L_ck.csr <- envelope(childcare_ck_ppp, Lest, nsim = 99, rank = 1, glocal=TRUE)
```


```{r}
plot(L_ck.csr, . - r ~ r, xlab="d", ylab="L(d)-r")
```

Tampines Planning Area

```{r}
L_tm = Lest(childcare_tm_ppp, correction = "Ripley")
plot(L_tm, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)", 
     xlim=c(0,1000))
```

Performing Complete Spatial Randomness Test
```{r}
L_tm.csr <- envelope(childcare_tm_ppp, Lest, nsim = 99, rank = 1, glocal=TRUE)
```


```{r}
plot(L_tm.csr, . - r ~ r, 
     xlab="d", ylab="L(d)-r", xlim=c(0,500))
```

End