---
title: "In-class Exercise 7"
author: "Kwek Ming Rong"
date: "20 Februrary 2023"
date-modified: "`r Sys.Date()`"
format: html
execute:
  eval: true
  echo: true
  warning: false
editor: visual
---

# Global and Local Measures of Spatial Autocorrelation - SFDEP

# Overview

In this hands-on exercise, you will learn how to compute Global and Local Measure of Spatial Autocorrelation (GLSA) by using **spdep** package. By the end to this hands-on exercise, you will be able to:

-   import geospatial data using appropriate function(s) of **sf** package,

-   import csv file using appropriate function of **readr** package,

-   perform relational join using appropriate join function of **dplyr** package,

-   compute Global Spatial Autocorrelation (GSA) statistics by using appropriate functions of **spdep** package,

    -   plot Moran scatterplot,

    -   compute and plot spatial correlogram using appropriate function of **spdep** package.

-   compute Local Indicator of Spatial Association (LISA) statistics for detecting clusters and outliers by using appropriate functions **spdep** package;

-   compute Getis-Ord's Gi-statistics for detecting hot spot or/and cold spot area by using appropriate functions of **spdep** package; and

-   to visualise the analysis output by using **tmap** package.

# 1. Installing and Loading R packages 

```{r}
pacman::p_load(sf, sfdep, tmap, tidyverse)
```

## 1.1 Importing the Geospatial Data

The code chunk below uses st_read() of sf package to import Hunan shapefile into R. The imported shapefile will be simple features Object of sf.

```{r}
hunan <- st_read(dsn = "Data/Geospatial", 
                 layer = "Hunan")
```

## 1.2 Importing CSV file into environment

Next, we will import Hunan_2012.csv into R by using read_csv() of readr package. The output is R dataframe class.

```{r}
hunan2012 <- read_csv("Data/Aspatial/Hunan_2012.csv")
```

## 1.3 Performing relational join

The code chunk below will be used to update the attribute table of hunan's SpatialPolygonsDataFrame with the attribute fields of hunan2012 dataframe. This is performed by using left_join() of dplyr package.

*In order to retain the geospatial properties, the left data frame must be the sf data.frame (i.e. hunan)*

```{r}
hunan_GDPPC <- left_join(hunan,hunan2012)%>%
  select(1:4, 7, 15)
```
# 2. Plotting a Choropleth Map

```{r}
tmap_mode("plot")
tm_shape(hunan_GDPPC) +
  tm_fill("GDPPC",
          style = "quantile", 
          palette = "Blues",
          title = "GDPPC") +
  tm_layout(main.title = "Distribution of GDP per captial by district",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45,
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar() + 
  tm_grid(alpha = 0.2)
```

# 3. Deriving the contiguity weights

## 3.1 Contiguity weights: Queen's method

```{r}
wm_q <- hunan_GDPPC %>%
  mutate(nb = st_contiguity(geometry),
         wt = st_weights(nb,
                         style = "W"),
         .before = 1)
```

## 3.2 Contiguity weights: Rook's method

```{r}
wm_r <- hunan_GDPPC %>%
  mutate(nb = st_contiguity(geometry),
  queen = FALSE,
  wt = st_weights(nb,
                  style = "W"),
  .before = 1)
```

## 3.3 Computing Global Moran I

```{r}
moranI <- global_moran(wm_q$GDPPC,
                        wm_q$nb,
                        wm_q$wt)
```

## 3.4 Performing Global Moran I's Test

```{r}
global_moran_test(wm_q$GDPPC,
                  wm_q$nb,
                  wm_q$wt)
```

## 3.5 Performing Global Moran's I permutation test

To ensure results stay the same when rendering every time

```{r}
set.seed(1234)
```

```{r}
global_moran_perm(wm_q$GDPPC,
                  wm_q$nb,
                  wm_q$wt,
                  nsim = 99)
```

If observation values are small, it is better to use a higher number of simulations

## 3.6 Computing local Moran's I

```{r}
lisa <- wm_q %>%
  mutate(local_moran = local_moran(
    GDPPC, nb, wt, nsim = 99),
    .before = 1) %>%
  unnest(local_moran)

lisa
```

Variable mean and pysal value should be the same. For take home exercise 2, stay with mean


