---
title: "In-class Exercise 6"
author: "Kwek Ming Rong"
date: "13 Februrary 2023"
date-modified: "`r Sys.Date()`"
format: html
execute:
  eval: true
  echo: true
  warning: false
editor: visual
---

# Installing and Loading the R packages

Three R packages will be used for this in class exercise, they are sf, sfdep and tidyverse.

```{r}
pacman::p_load(sf, sfdep, tidyverse, tmap)
```

# 1. The Data

For the purpose of this in class exercise, the Hunan data sets will be used. There are two data sets in this use case, they are

-   Hunan county boundary layer. This is a geospatial data set in ESRI shapefile format.

-   Hunan_2012.csv: This csv file contains selected Hunan's local development indicators in 2012.

## 1.1 Importing the Geospatial Data

The code chunk below uses st_read() of sf package to import Hunan shapefile into R. The imported shapefile will be simple features Object of sf.

```{r}
hunan <- st_read(dsn = "Data/Geospatial", 
                 layer = "Hunan")
```

## 1.2 Importing CSV file into environment

Next, we will import Hunan_2012.csv into R by using read_csv() of readr package. The output is R dataframe class.

```{r}
hunan2012 <- read_csv("Data/Aspatial/Hunan_2012.csv")
```

## 1.3 Performing relational join

The code chunk below will be used to update the attribute table of hunan’s SpatialPolygonsDataFrame with the attribute fields of hunan2012 dataframe. This is performed by using left_join() of dplyr package.

*In order to retain the geospatial properties, the left data frame must be the sf data.frame (i.e. hunan)*

```{r}
hunan_GDPPC <- left_join(hunan,hunan2012)%>%
  select(1:4, 7, 15)
```

# 2. Plotting a Choropleth Map

```{r}
basemap <- tm_shape(hunan_GDPPC) +
  tm_polygons() +
  tm_text("NAME_3", size=0.5)

gdppc <- qtm(hunan_GDPPC, "GDPPC")
tmap_arrange(basemap, gdppc, asp=1, ncol=2)
```